Foreign direct investment, corporate social responsibility, and malaria control in Mozambique - trends, risks, and opportunities

### 2017-07-11 paper.Rmd
incorporating elisas fmi data

>>>>

### Sources of FDI in Mozambique

```{r}
# Define function for reading fdi data from imf
read_imf <- function(file){
  srcs <- readxl::read_excel(file, skip = 4)
  # Rename columns
  srcs <- srcs[,1:3]
  srcs <- srcs[2:nrow(srcs),]
  names(srcs) <- c('from_country', 'inward', 'outward')
  # Make numberic
  srcs <- data.frame(srcs)
  for (j in 2:3){
    this_column <- as.numeric(as.character(gsub('.', '', srcs[,j], fixed = TRUE)))
    srcs[,j] <- this_column
  }
  srcs <- srcs %>%
    filter(from_country != 'World')

  # Get amount by inward
  x <- srcs %>%
    arrange(desc(inward)) %>%
    filter(!is.na(inward) & inward > 0) %>%
    mutate(from_country = ifelse(from_country == 'China, P.R.: Mainland',
                                 'China',
                                 ifelse(from_country == 'United Arab Emirates',
                                        'U.A.E.',
                                        ifelse(from_country == 'United Kingdom',
                                               'U.K.',
                                               from_country)))) #%>%
    # mutate(from_country = factor(from_country, levels = from_country))
  return(x)
}

years <- 2009:2015
results_list <- list()
for (i in 1:length(years)){
  this_year <- years[i]
  this_data <- read_imf(paste0('data/imf_inward_direct_investment_positions/', this_year, '.xls')) %>%
    mutate(year = this_year)
  results_list[[i]] <- this_data
}
x <- bind_rows(results_list)
```

According to the IMF, `r x$from_country[x$year == 2015 & !is.na(x$inward)]` countries had foreign direct investment in Mozambique as of the end of 2015. The below chart shows the top 10 countries ranked by amount invested in Mozambique.

```{r}
last_year <- x %>%
  filter(year == 2015)
last_year <- last_year %>% 
  arrange(desc(inward)) %>%
  mutate(from_country = factor(from_country, levels = as.character(from_country)))
last_year <- last_year[1:10,]
ggplot(data = last_year,
       aes(x = from_country,
           y = inward)) +
  geom_bar(stat = 'identity',
           alpha = 0.6,
           fill = 'darkblue') +
  labs(x = 'Source of investment',
       y = '2015 USD, Millions',
       title = 'Sources of FDI in Mozambique',
       subtitle = 'By investor country') +
  theme_bw() +
    theme(axis.text.x = element_text(angle = 90)) 
```

What stands out is the presence of many "non-Western" countries in the top 10. Though data are not available over time (ie, we cannot make temporal comparisons), the landscape of investment 20 years prior would have been both (a) far smaller and (b) primarily composed of western countries, especially given that Mozambique emerged so late from civil war (1994).

If we examine the same IMF data over time, we can see that even in the short period for which data is available (2009 to 2015) there is a general upward trend, but that the trend is steepest in the last two years from China:

```{r}
cols <- colorRampPalette(brewer.pal(n = 9, name = 'Spectral'))(nrow(last_year))
cols[3] <- 'black'
ggplot(data = x %>%
         filter(from_country %in% last_year$from_country),
       aes(x = year,
           y = inward,
           group = from_country,
           color = from_country)) +
  geom_line() +
  scale_color_manual(name = 'Country',
                     values = cols) +
labs(x = 'Source of investment',
       y = '2015 USD, Millions',
       title = 'Sources of FDI in Mozambique over time',
       subtitle = 'By investor country') +
  theme_bw() 
```

### 2017-07-11 paper.tex
incorporating elisas fmi data

@@

@@

>>>>

\subsubsection{Sources of FDI in
Mozambique}\label{sources-of-fdi-in-mozambique}

According to the IMF, U.K., Zambia, China, Canada, Turkey, U.A.E.,
Romania, India, Bulgaria, Lebanon, Argentina, Singapore, Netherlands,
Spain, Norway, Angola, United States, China, P.R.: Macao, Denmark,
Botswana, Serbia, Republic of, Guadeloupe, Swaziland, Bermuda, Panama,
Japan, Switzerland, Ireland, Malaysia, Russian Federation, Nigeria,
Uruguay, Mauritius, Italy, Burundi, Kuwait, Guinea-Bissau, Australia,
Zimbabwe, Togo, China, P.R.: Hong Kong, Virgin Islands, British,
Portugal, Iceland, Austria, Sweden, Malta, Isle of Man, Suriname,
Tanzania, Brazil, Korea, Republic of, Bahamas, The, Saudi Arabia, Ghana,
Seychelles, Reunion, Kenya, South Africa, France, Congo, Democratic
Republic of, Germany, Uganda, Thailand, Slovak Republic, Estonia, Puerto
Rico, Cyprus, Indonesia countries had foreign direct investment in
Mozambique as of the end of 2015. The below chart shows the top 10
countries ranked by amount invested in Mozambique.

What stands out is the presence of many ``non-Western'' countries in the
top 10. Though data are not available over time (ie, we cannot make
temporal comparisons), the landscape of investment 20 years prior would
have been both (a) far smaller and (b) primarily composed of western
countries, especially given that Mozambique emerged so late from civil
war (1994).

If we examine the same IMF data over time, we can see that even in the
short period for which data is available (2009 to 2015) there is a
general upward trend, but that the trend is steepest in the last two
years from China:

@@

\begin{center}\includegraphics{paper_files/figure-latex/unnamed-chunk-13-1} \end{center}

@@

\begin{center}\includegraphics{paper_files/figure-latex/unnamed-chunk-18-1} \end{center}
